---
title: "Tidy Tuesday Exercise"
---

### Load Data

```{r}
# Load necessary libraries
library(tidyverse)
library(caret)
library(tidymodels)
library(ggplot2)

# Load the CSV files
ewf_matches <- read_csv("data/ewf_matches.csv")
ewf_appearances <- read_csv("data/ewf_appearances.csv")
ewf_standings <- read_csv("data/ewf_standings.csv")

# Summarize the datasets
summary(ewf_matches)
summary(ewf_appearances)
summary(ewf_standings)

# Number of matches per season
matches_per_season <- ewf_matches %>%
  group_by(season) %>%
  summarise(total_matches = n())

# Plot the number of matches per season
ggplot(matches_per_season, aes(x = season, y = total_matches)) +
  geom_bar(stat = "identity") +
  theme_minimal() +
  labs(title = "Number of Matches per Season", x = "Season", y = "Total Matches")

# Average attendance per season
average_attendance <- ewf_matches %>%
  mutate(attendance = as.numeric(gsub(",", "", attendance))) %>%
  group_by(season) %>%
  summarise(average_attendance = mean(attendance, na.rm = TRUE))

# Plot the average attendance per season
ggplot(average_attendance, aes(x = season, y = average_attendance)) +
  geom_line() +
  geom_point() +
  theme_minimal() +
  labs(title = "Average Attendance per Season", x = "Season", y = "Average Attendance")

# Number of goals scored by each team per season
goals_per_team <- ewf_standings %>%
  group_by(season, team_name) %>%
  summarise(total_goals = sum(goals_for))

# Plot the goals scored by each team per season
ggplot(goals_per_team, aes(x = team_name, y = total_goals, fill = season)) +
  geom_bar(stat = "identity", position = "dodge") +
  theme_minimal() +
  labs(title = "Goals Scored by Each Team per Season", x = "Team", y = "Total Goals") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

# Distribution of match results
match_results <- ewf_matches %>%
  group_by(result) %>%
  summarise(count = n())

# Plot the distribution of match results
ggplot(match_results, aes(x = result, y = count, fill = result)) +
  geom_bar(stat = "identity") +
  theme_minimal() +
  labs(title = "Distribution of Match Results", x = "Result", y = "Count")

# Save the plots to files
ggsave("matches_per_season.png", plot = last_plot(), width = 8, height = 5)
ggsave("average_attendance_per_season.png", plot = last_plot(), width = 8, height = 5)
ggsave("goals_per_team.png", plot = last_plot(), width = 8, height = 5)
ggsave("match_results_distribution.png", plot = last_plot(), width = 8, height = 5)

# Print the summaries and head of each dataset
print("Summary of ewf_matches:")
print(summary(ewf_matches))

print("Summary of ewf_appearances:")
print(summary(ewf_appearances))

print("Summary of ewf_standings:")
print(summary(ewf_standings))

print("Head of ewf_matches:")
print(head(ewf_matches))

print("Head of ewf_appearances:")
print(head(ewf_appearances))

print("Head of ewf_standings:")
print(head(ewf_standings))

```

### Question:

How has team performance evolved over the seasons in the Women's Super League?

### Hypothesis:

Teams that have higher average attendance tend to perform better in terms of points and goal difference.

```{r}


# Data Preparation
ewf_matches <- ewf_matches %>%
  mutate(attendance = as.numeric(gsub(",", "", attendance))) %>%
  group_by(season, home_team_id) %>%
  summarise(avg_attendance = mean(attendance, na.rm = TRUE), .groups = 'drop')

# Merge datasets
team_performance <- ewf_standings %>%
  left_join(ewf_matches, by = c("season", "team_id" = "home_team_id"))

# Handle missing values
team_performance <- team_performance %>%
  mutate(avg_attendance = ifelse(is.na(avg_attendance), mean(avg_attendance, na.rm = TRUE), avg_attendance))

# Select relevant columns and convert data types if necessary
team_performance <- team_performance %>%
  select(season, team_name, points, goals_for, goals_against, wins, losses, draws, avg_attendance)

# Splitting the data into training and test sets
set.seed(123)  # For reproducibility
train_index <- createDataPartition(team_performance$points, p = 0.7, list = FALSE)
train_data <- team_performance[train_index, ]
test_data <- team_performance[-train_index, ]

# Set up the resampling
set.seed(123)
cv_folds <- vfold_cv(train_data, v = 5)

# Define Models
linear_reg_model <- linear_reg() %>%
  set_engine("lm") %>%
  set_mode("regression")

rf_model <- rand_forest() %>%
  set_engine("ranger") %>%
  set_mode("regression")

svm_model <- svm_rbf() %>%
  set_engine("kernlab") %>%
  set_mode("regression")

# Create Recipe
team_recipe <- recipe(points ~ goals_for + goals_against + wins + losses + draws + avg_attendance, data = train_data) %>%
  step_normalize(all_predictors())

# Set up Workflows
linear_reg_workflow <- workflow() %>%
  add_model(linear_reg_model) %>%
  add_recipe(team_recipe)

rf_workflow <- workflow() %>%
  add_model(rf_model) %>%
  add_recipe(team_recipe)

svm_workflow <- workflow() %>%
  add_model(svm_model) %>%
  add_recipe(team_recipe)

# Train Models Using Cross-Validation
set.seed(123)
linear_reg_results <- fit_resamples(
  linear_reg_workflow,
  resamples = cv_folds,
  metrics = metric_set(rmse, rsq)
)

set.seed(123)
rf_results <- fit_resamples(
  rf_workflow,
  resamples = cv_folds,
  metrics = metric_set(rmse, rsq)
)

set.seed(123)
svm_results <- fit_resamples(
  svm_workflow,
  resamples = cv_folds,
  metrics = metric_set(rmse, rsq)
)

# Collect and Print Metrics
linear_reg_metrics <- collect_metrics(linear_reg_results)
rf_metrics <- collect_metrics(rf_results)
svm_metrics <- collect_metrics(svm_results)

print("Linear Regression Metrics")
print(linear_reg_metrics)

print("Random Forest Metrics")
print(rf_metrics)

print("SVM Metrics")
print(svm_metrics)

# Visualize Residuals and Predictions
plot_residuals <- function(results, title) {
  collect_predictions(results) %>%
    ggplot(aes(x = .pred, y = .pred - points)) +
    geom_point(alpha = 0.5) +
    geom_hline(yintercept = 0, linetype = "dashed") +
    labs(title = title, x = "Predicted", y = "Residuals") +
    theme_minimal()
}

# Function to plot residuals
plot_residuals <- function(results, title) {
  results %>%
    collect_predictions() %>%
    ggplot(aes(x = .pred, y = .pred - points)) +
    geom_point(alpha = 0.5) +
    geom_hline(yintercept = 0, linetype = "dashed") +
    labs(title = title, x = "Predicted", y = "Residuals") +
    theme_minimal()
}




```

### Model Fitting

```{r}
# Linear Regression Model
linear_reg_model <- linear_reg() %>%
  set_engine("lm") %>%
  set_mode("regression")

# Random Forest Model
rf_model <- rand_forest() %>%
  set_engine("ranger") %>%
  set_mode("regression")

# Support Vector Machine Model
svm_model <- svm_rbf() %>%
  set_engine("kernlab") %>%
  set_mode("regression")
# Recipe for preprocessing
team_recipe <- recipe(points ~ goals_for + goals_against + wins + losses + draws + avg_attendance, data = train_data) %>%
  step_normalize(all_predictors())
# Set up workflows
linear_reg_workflow <- workflow() %>%
  add_model(linear_reg_model) %>%
  add_recipe(team_recipe)

rf_workflow <- workflow() %>%
  add_model(rf_model) %>%
  add_recipe(team_recipe)

svm_workflow <- workflow() %>%
  add_model(svm_model) %>%
  add_recipe(team_recipe)
# Train and evaluate Linear Regression Model
set.seed(123)
linear_reg_results <- fit_resamples(
  linear_reg_workflow,
  resamples = cv_folds,
  metrics = metric_set(rmse, rsq)
)

# Train and evaluate Random Forest Model
set.seed(123)
rf_results <- fit_resamples(
  rf_workflow,
  resamples = cv_folds,
  metrics = metric_set(rmse, rsq)
)

# Train and evaluate SVM Model
set.seed(123)
svm_results <- fit_resamples(
  svm_workflow,
  resamples = cv_folds,
  metrics = metric_set(rmse, rsq)
)
# Collect metrics
linear_reg_metrics <- collect_metrics(linear_reg_results)
rf_metrics <- collect_metrics(rf_results)
svm_metrics <- collect_metrics(svm_results)

# Print metrics
print("Linear Regression Metrics")
print(linear_reg_metrics)

print("Random Forest Metrics")
print(rf_metrics)

print("SVM Metrics")
print(svm_metrics)


```

### Model Selection

While the Linear Regression model has the best performance in terms of both RMSE and R-squared, the choice of the best model should also consider factors beyond just performance metrics. Here’s a detailed rationale for selecting the best model:

### Linear Regression

#### Pros:

1.  **Performance**: The Linear Regression model has the lowest RMSE and the highest R-squared, indicating it has the best predictive performance among the three models.

2.  **Interpretability**: Linear Regression is highly interpretable, allowing for clear insights into how each predictor variable affects the outcome. This is crucial for understanding the factors that contribute to team performance in women's football.

3.  **Simplicity**: Linear Regression is simple and computationally efficient, making it easy to implement and understand.

#### Cons:

1.  **Assumptions**: Linear Regression relies on several assumptions (linearity, homoscedasticity, independence, and normality of residuals), which might not always hold true in real-world data.

### Random Forest

#### Pros:

1.  **Performance**: The Random Forest model also performs well, with a reasonably low RMSE and a high R-squared. It handles non-linear relationships and interactions between variables better than Linear Regression.

2.  **Robustness**: Random Forest is robust to outliers and can handle a large number of predictors and complex interactions.

#### Cons:

1.  **Interpretability**: Random Forest models are less interpretable compared to Linear Regression. Understanding the contribution of each predictor to the outcome can be more challenging.

2.  **Computational Cost**: Random Forests are more computationally intensive, especially with a large number of trees.

### Support Vector Machine (SVM)

#### Pros:

1.  **Complex Relationships**: SVMs can capture complex, non-linear relationships between the predictors and the outcome.

#### Cons:

1.  **Performance**: The SVM model has the highest RMSE and the lowest R-squared among the three models, indicating it performs the worst in terms of predictive accuracy.

2.  **Interpretability**: SVMs are even less interpretable than Random Forests, making it difficult to understand the model's decisions.

3.  **Computational Cost**: SVMs can be computationally expensive, especially with large datasets.

### Conclusion

Given the above considerations, I would select the **Linear Regression** model as the overall best model. The primary reasons for this choice are its superior performance in terms of RMSE and R-squared, its simplicity, and its high interpretability, which aligns well with the goal of understanding the factors contributing to team performance in women's football.

### Final Model Choice: Linear Regression

#### Explanation:

-   **Performance**: Best predictive performance with the lowest RMSE and highest R-squared.

-   **Interpretability**: Allows for clear insights into the effect of each predictor on the outcome.

-   **Simplicity**: Easy to implement, understand, and communicate the results.

This choice ensures that the analysis remains transparent and actionable, providing valuable insights into the factors driving team performance while maintaining high predictive accuracy.

### Test Data Validation

```{r}
# Load necessary libraries
library(tidyverse)
library(tidymodels)

# Load the CSV files
ewf_matches <- read_csv("data/ewf_matches.csv")
ewf_appearances <- read_csv("data/ewf_appearances.csv")
ewf_standings <- read_csv("data/ewf_standings.csv")

# Data Preparation
ewf_matches <- ewf_matches %>%
  mutate(attendance = as.numeric(gsub(",", "", attendance))) %>%
  group_by(season, home_team_id) %>%
  summarise(avg_attendance = mean(attendance, na.rm = TRUE), .groups = 'drop')

# Merge datasets
team_performance <- ewf_standings %>%
  left_join(ewf_matches, by = c("season", "team_id" = "home_team_id"))

# Handle missing values
team_performance <- team_performance %>%
  mutate(avg_attendance = ifelse(is.na(avg_attendance), mean(avg_attendance, na.rm = TRUE), avg_attendance))

# Select relevant columns and convert data types if necessary
team_performance <- team_performance %>%
  select(season, team_name, points, goals_for, goals_against, wins, losses, draws, avg_attendance)

# Splitting the data into training and test sets
set.seed(123)  # For reproducibility
train_index <- createDataPartition(team_performance$points, p = 0.7, list = FALSE)
train_data <- team_performance[train_index, ]
test_data <- team_performance[-train_index, ]
# Define the Linear Regression Model
linear_reg_model <- linear_reg() %>%
  set_engine("lm") %>%
  set_mode("regression")

# Create Recipe
team_recipe <- recipe(points ~ goals_for + goals_against + wins + losses + draws + avg_attendance, data = train_data) %>%
  step_normalize(all_predictors())

# Set up Workflow
linear_reg_workflow <- workflow() %>%
  add_model(linear_reg_model) %>%
  add_recipe(team_recipe)

# Fit the Model
linear_reg_fit <- linear_reg_workflow %>%
  fit(data = train_data)
# Make Predictions on Test Data
test_predictions <- predict(linear_reg_fit, new_data = test_data) %>%
  bind_cols(test_data)

# Evaluate Model Performance
test_metrics <- test_predictions %>%
  metrics(truth = points, estimate = .pred)

# Print Test Metrics
print(test_metrics)

# Plot Residuals
test_residuals_plot <- test_predictions %>%
  ggplot(aes(x = .pred, y = .pred - points)) +
  geom_point(alpha = 0.5) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  labs(title = "Linear Regression Residuals on Test Data", x = "Predicted", y = "Residuals") +
  theme_minimal()

print(test_residuals_plot)


```
